---
title: "extract_data_to_subset-topography"
author: "Adriana Uscanga"
date: "2023-05-25"
output: html_document
---

## Description
Code to clip topographic data to the extent of the sample areas (10 burned and 10 unburned polygons) and extract raster values to points from sampling grid. Polygons and points shapefiles can be found here: https://drive.google.com/drive/u/2/folders/1K9MRkhiBOF70WU27HRc5ZOS-SsOYjG8j
I downloaded topographic data from the workshop data_sandbox folder through cyverse: /iplant/home/shared/earthlab/Macrosystems_workshop/data_sandbox/topography_southern_rockies.tif
The topography_southern_rockies.tif file does not fully cover the sample areas, thus, I also downloaded a dem (NASA SRTM-30m using google earth engine) to clip and extract data (see With DEM section below).

### Load packages and set paths
```{r}

### Load packages
library(tidyverse)
library(sf)
library(sp)
library(raster)

### Set path for files -- I replaced this to set the path to my local computer
### file path for the subset of polygons and grid points
path_subset <- "/Users/adrianauscanga/Documents/Documents/recovery-group/adriana_scratch/subset_polygons"

### file path for the raster (here, I'm using Landfire's EVT as an example)
path_raster <- "/Users/adrianauscanga/Documents/Documents/recovery-group/adriana_scratch"

### file path for cropped datasets
path_saveto <- "/Users/adrianauscanga/Documents/Documents/recovery-group/adriana_scratch"
```

## With topography_southern_rockies.tif
### Open data layers and harmonize projections

```{r}
### Open raster
topo <- raster(file.path(path_raster, "topography_southern_rockies.tif"))
names(topo) <- "elevation"
### All CONUS landfire data is projected into USA Contiguous Albers Equal Area Conic (EPSG:5070)

### Open polygons and transform to EPSG of raster layer
subset_polys <- st_read(file.path(path_subset, "/subset_polys.shp")) %>%
  
  ### transform projection to match raster
  st_transform(., crs = crs(topo))

### Open sample grid and transform to EPSG of raster layer
subset_points <- st_read(file.path(path_subset, "/subset_points.shp")) %>%
  
  ### transform projection to match raster
  st_transform(., crs = crs(topo))
```

### Subset data to subset of polygons
```{r}
## POLYGONS
### Crop raster to subset of polygons
crop_raster <- mask(crop(topo, 
                         extent(subset_polys)), 
                    subset_polys)

## POINTS
### Make the points sf an sp object
points_sp <- as(subset_points, 'Spatial')

### Extract raster values to grid samples
extract_raster <- raster::extract(topo,
                                  points_sp,
                                  df = TRUE,
                                  sp = TRUE)

### Convert back to sf
extract_raster <- extract_raster %>%
  st_as_sf()

## Write out files
### write out raster cropped to polygons
writeRaster(crop_raster,
            filename = file.path(path_saveto, "topo_cropped.tif"))

### write out shp of raster values extracted to points
st_write(extract_raster,
         file.path(path_saveto, "topo_points.shp"))
```

## With DEM

### Open data layers and harmonize projections

```{r}
### Open raster
elevation <- raster(file.path(path_raster, "elevation_southern_rockies.tif"))

### Open polygons and transform to EPSG of raster layer
subset_polys <- st_read(file.path(path_subset, "/subset_polys.shp")) %>%
  
  ### transform projection to match raster
  st_transform(., crs = crs(elevation))

### Open sample grid and transform to EPSG of raster layer
subset_points <- st_read(file.path(path_subset, "/subset_points.shp")) %>%
  
  ### transform projection to match raster
  st_transform(., crs = crs(elevation))
```

### Subset data to subset of polygons
```{r}
## POLYGONS
### Crop raster to subset of polygons
crop_raster <- mask(crop(elevation, 
                         extent(subset_polys)), 
                    subset_polys)

## POINTS
### Make the points sf an sp object
points_sp <- as(subset_points, 'Spatial')

### Extract raster values to grid samples
extract_raster <- raster::extract(elevation,
                                  points_sp,
                                  df = TRUE,
                                  sp = TRUE)

### Convert back to sf
extract_raster <- extract_raster %>%
  st_as_sf()

## Write out files
### write out raster cropped to polygons
writeRaster(crop_raster,
            filename = file.path(path_saveto, "elevation_cropped.tif"))

### write out shp of raster values extracted to points
st_write(extract_raster,
         file.path(path_saveto, "elevation_points.shp"))
```
