---
title: "extract_data_to_subset - NLCD"
author: "author"
output: html_document
date: "2023-04-27"
---

## Description

Code to clip NLCD rasters to the extent of the sample areas and extract raster values to points from sampling grid, based on code by Katherine Siegel.

## R:
### Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### Load packages
library(tidyverse)
library(sf)
library(sp)
library(raster)
library(here)
#install.packages("devtools")
#devtools::install_github("ropensci/FedData")
library(FedData)
library(rgdal)#

### Set path for files -- replace this in your script!

### file path for the subset of polygons and grid points
path_subset <- here("recovery-group", "shapefiles")

### file path for the raster (here, I'm using Landfire's EVT as an example)
path_raster <- here("recovery-group", "tifs")

### file path for cropped datasets
path_saveto <- here("recovery-group", "nlcd_cropped_datasets")
```

#---------------------------

#Way 1: (creates a WIDE dataset)

Read in tiffs (downloaded from https://www.mrlc.gov/viewer/)

```{r}
nlcd_2001 <- raster(file.path(path_raster, "NLCD_2001_Land_Cover_L48.tiff"))
nlcd_2004 <- raster(file.path(path_raster, "NLCD_2004_Land_Cover_L48.tiff"))
nlcd_2006 <- raster(file.path(path_raster, "NLCD_2006_Land_Cover_L48.tiff"))
nlcd_2008 <- raster(file.path(path_raster, "NLCD_2008_Land_Cover_L48.tiff"))
nlcd_2011 <- raster(file.path(path_raster, "NLCD_2011_Land_Cover_L48.tiff"))
nlcd_2013 <- raster(file.path(path_raster, "NLCD_2013_Land_Cover_L48.tiff"))
nlcd_2016 <- raster(file.path(path_raster, "NLCD_2016_Land_Cover_L48.tiff"))
nlcd_2019 <- raster(file.path(path_raster, "NLCD_2019_Land_Cover_L48.tiff"))

nlcd_allyears <- raster::stack(nlcd_2001, nlcd_2004, nlcd_2006, nlcd_2008, nlcd_2011, nlcd_2013, nlcd_2016, nlcd_2019)

plot(nlcd_allyears)
```


```{r}
my_crs <- crs(nlcd_allyears)

### Open polygons and transform to EPSG of raster layer
subset_polys <- st_read(file.path(path_subset, "subset_polys.shp")) %>%
  ### transform projection to match raster
  st_transform(., crs = my_crs)
subset_polys

### Open sample grid and transform to EPSG of raster layer
subset_points <- st_read(file.path(path_subset, "subset_points.shp")) %>%
  
  ### transform projection to match raster
  st_transform(., crs = my_crs)
```


### Subset data to subset of polygons
```{r, warning = F}
## POLYGONS
### Crop raster to subset of polygons
crop_raster <- mask(crop(nlcd_allyears, 
                         extent(subset_polys)), 
                    subset_polys)

## POINTS
### Make the points sf an sp object
points_sp <- as(subset_points, 'Spatial')

### Extract raster values to grid samples
extract_raster_layer <- raster::extract(nlcd_allyears,
                                  points_sp,
                                  df = TRUE,
                                  sp = TRUE)

### Convert back to sf
extract_raster <- extract_raster_layer %>%
  st_as_sf() %>% 
  rename(nlcd_2001 = Layer_1.1,
         nlcd_2004 = Layer_1.2,
         nlcd_2006 = Layer_1.3,
         nlcd_2008 = Layer_1.4,
         nlcd_2011 = Layer_1.5,
         nlcd_2013 = Layer_1.6,
         nlcd_2016 = Layer_1.7,
         nlcd_2018 = Layer_1.8)
```


```{r, warning = F}
#add legend: 
legend<-pal_nlcd() %>% 
  as.data.frame() %>% 
  mutate(nlcd_code = ID)
legend

write.csv(legend, here("nlcd_metadata.csv"))

extract_raster_legend <- merge(extract_raster, legend, by = c("nlcd_code"))
```


```{r, warning = F}
## Write out files
### write out raster cropped to polygons
writeRaster(crop_raster,filename = file.path(path_saveto, "nlcd_rast_cropped_wide.tif"), overwrite=TRUE)

### write out shp of raster values extracted to points
st_write(extract_raster, file.path(path_saveto, "nlcd_rast_points_wide.shp"), append = FALSE)
```

#---------------------------

#Way 2: 

using get_nlcd() function, creates a LONG dataset. 

Do the following for each year (could probably make easier with a loop, but I just did year by year and then stacked with raster::stack() at the end before extracting to polygons/points. Only 2001 shown here. 

###Get 2001 nlcd data: 

```{r}
nlcd_2001 <-
  get_nlcd(
    template = subset_polys,
    label = "2001",
    year = 2001
  )
```
```{r}
###NLCD is in WGS84 Coordinate Systems (EPSG: 4326), but we can just define the CRS this way without having to set the code:
my_crs <- crs(nlcd_2001)

### Open polygons and transform to EPSG of raster layer
subset_polys <- st_read(file.path(path_subset, "subset_polys.shp")) %>%
  ### transform projection to match raster
  st_transform(., crs = my_crs)

### Open sample grid and transform to EPSG of raster layer
subset_points <- st_read(file.path(path_subset, "subset_points.shp")) %>%
  
  ### transform projection to match raster
  st_transform(., crs = my_crs)
```

### Subset data to subset of polygons
```{r}
## POLYGONS
### Crop raster to subset of polygons
crop_2001 <- mask(crop(nlcd_2001, 
                         extent(subset_polys)), 
                    subset_polys)

## POINTS
### Make the points sf an sp object
points_sp <- as(subset_points, 'Spatial')

### Extract raster values to grid samples
extract_raster_layer <- raster::extract(nlcd_2001,
                                  points_sp,
                                  df = TRUE,
                                  sp = TRUE)

### Convert back to sf
extract_raster_2001 <- extract_raster_layer %>%
  st_as_sf() %>% 
  mutate(nlcd_code = Class) %>% 
  dplyr::select(-Class)

#add legend: 
legend<-pal_nlcd() %>% 
  as.data.frame() %>% 
  mutate(nlcd_code = ID)
legend

extract_raster_legend_2001 <- merge(extract_raster_2001, legend, by = c("nlcd_code")) %>% 
  mutate(nlcd_class = Class, 
         nlcd_color = Color, 
         nlcd_class_description = Description, 
         nlcd_year = 2001) %>% 
  dplyr::select(-ID, -Class, -Color, -Description) 

## Write out files
### write out raster cropped to polygons
writeRaster(crop_2001,filename = file.path(path_saveto, "nlcd_rast_2001_cropped.tif"), overwrite=TRUE)

### write out shp of raster values extracted to points
st_write(extract_raster_legend_2001, file.path(path_saveto, "nlcd_rast_points_2001.shp"), append = F)
```

##Put all together: 

Polygons: 

```{r}
nlcd_all <- raster::stack(crop_2019, 
                         crop_2016,
                         crop_2013,
                         crop_2011,
                         crop_2008,
                         crop_2006,
                         crop_2004,
                         crop_2001
                         )

names(nlcd_all) <- c("2019", "2016", "2013", "2011", "2008", "2006", "2004", "2001")

plot(nlcd_all)

## Write out files
### write out raster cropped to polygons
writeRaster(nlcd_all,filename = file.path(path_saveto, "nlcd_rast_alldates_cropped.tif"), overwrite=TRUE)

# read from disk, check
r2 <- terra::rast(file.path(path_saveto, "nlcd_rast_alldates_cropped.tif"))
r2

plot(r2)
```

Points: 

```{r, warning = F}
nlcd_raster_all <- bind_rows(extract_raster_legend_2019,
                             extract_raster_legend_2016,
                             extract_raster_legend_2013,
                             extract_raster_legend_2011,
                             extract_raster_legend_2008,
                             extract_raster_legend_2006,
                             extract_raster_legend_2004,
                             extract_raster_legend_2001) %>% 
  janitor::clean_names()

### write out shp of raster values extracted to points
st_write(nlcd_raster_all, file.path(path_saveto, "nlcd_points_alldates.shp"), append = F)
```

#---------------------------

#Change from 2001 to 2019

```{r}
nlcd_change <- raster(file.path(path_raster, "NLCD_2001_2019_change_index_L48.tiff"))

plot(nlcd_change)
```

```{r}
my_crs <- crs(nlcd_change)

### Open polygons and transform to EPSG of raster layer
subset_polys <- st_read(file.path(path_subset, "subset_polys.shp")) %>%
  ### transform projection to match raster
  st_transform(., crs = my_crs)
subset_polys

### Open sample grid and transform to EPSG of raster layer
subset_points <- st_read(file.path(path_subset, "subset_points.shp")) %>%
  
  ### transform projection to match raster
  st_transform(., crs = my_crs)
```

### Subset data to subset of polygons
```{r, warning = F}
## POLYGONS
### Crop raster to subset of polygons
crop_raster <- mask(crop(nlcd_change, 
                         extent(subset_polys)), 
                    subset_polys)

## POINTS
### Make the points sf an sp object
points_sp <- as(subset_points, 'Spatial')

### Extract raster values to grid samples
extract_raster_layer <- raster::extract(nlcd_change,
                                  points_sp,
                                  df = TRUE,
                                  sp = TRUE)

### Convert back to sf
extract_raster <- extract_raster_layer %>%
  st_as_sf() %>% 
  rename(nlcd_change_code = Layer_1) 
```

```{r, warning = F}
## Write out files
### write out raster cropped to polygons
writeRaster(crop_raster,filename = file.path(path_saveto, "nlcd_rast_cropped_nlcd_2001_2019_change_index.tif"), overwrite=TRUE)

### write out shp of raster values extracted to points
st_write(extract_raster, file.path(path_saveto, "nlcd_rast_points_nlcd_2001_2019_change_index.shp"), append = FALSE)
```


